---
title: "European Social Survey: Exploratory data analysis"
output: html_notebook
---

## Data

```{r echo=FALSE}
library(tidyverse)
library(essurvey)
library(labelled)
library(rstan)
set_email('trubetskoy.vasa@gmail.com')
```

Round 8 contains questions refering to climate. We can extract these.

```{r}
round_8 <- import_rounds(8)

base_variables <- c('idno', 'cntry')

interesting_variables <- c('eduyrs')

questions <- var_label(round_8) %>%
  as_data_frame() %>%
  gather(key = 'variable_name', value = 'description')

climate_questions <- questions %>%
  filter(str_detect(description, 'climate'))

responses <- round_8 %>%
  select(one_of(c(base_variables, climate_questions$variable_name, interesting_variables)))

# is this the right thing to do?
responses <- recode_missings(responses, c("Don't know", "Refusal"))

country_summaries <- responses %>%
  group_by(cntry) %>%
  summarise(mean_inctxff = mean(inctxff, na.rm = T))
```

## Mapping

Starting point:
https://gist.github.com/stared/fbca436c885c430a314a

```{r}
library(rgdal)
library(doBy)

eurMap <- readOGR('data/NUTS_RG_01M_2016_3035.shp')
eurMap <- subset(eurMap, nchar(as.character(NUTS_ID)) == 2)

eurMapDf <- fortify(eurMap, region='NUTS_ID')

# merge map and data
eurMapDataDf <- merge(eurMapDf, country_summaries, by.x="id", by.y="cntry")
# sort, so that polygons are drawn correctly
eurMapDataDf <- eurMapDataDf[order(eurMapDataDf$order),]

# limit data to main Europe
eurMapDataDf <- subset(eurMapDataDf, long > 2e6 & long < 6e6 & lat > 1e6 & lat < 6e6)

# add text; instead of mean I do middle (not to be to biased towards detailed coastlines)
middle = function (x) {
  (max(x) + min(x)) / 2
}  
txtVal <- summaryBy(long + lat + mean_inctxff ~ id, data=eurMapDataDf, FUN=middle, keep.names=T)

p <- ggplot(data=eurMapDataDf) +
  geom_polygon(aes(x=long, y=lat, group=group, fill=mean_inctxff)) +
  geom_path(aes(x=long, y=lat, group=group), color='black', alpha=.5) +
  geom_text(aes(x=long, y=lat, label=sprintf("%.1f", mean_inctxff)), data=txtVal, col="gray", cex=3) +
  scale_fill_gradient2() +
  theme_void() +
  coord_equal() + 
  ggtitle('"Increasing taxes on fossil fuels, such as oil, gas and coal."\n  Scale: 1=strongly in favor, 5=strongly against')
p
```

Weirdly the UK is missing on this map.

## Modeling

### Questions

How does income affect climate views?

How does level of education affect climate views?

How do climate views differ by country?

### Model 01: linear everything

How unrealistic! We can take a look anyway:

```{r}
preprocessed_data <- responses %>%
  select(c(eduyrs, inctxff)) %>%
  drop_na() %>%
  mutate(x = scale(eduyrs) %>% as.vector(), y = scale(inctxff) %>% as.vector())

ggplot(preprocessed_data, aes(x=x, y=y)) +
  geom_point() +
  geom_smooth(method='lm',formula=y~x) +
  theme_bw()
```

Yikes. Lets fit this thing anyway.

```{r}
stan_data <- preprocessed_data %>%
  select(c(x, y)) %>%
  as.list()

stan_data[['N']] <- length(stan_data[['y']])

edu_inctxf_fit <-stan(file = '../models/linear_education_climate.stan',
  data = stan_data,
  iter = 2000,
  warmup = 200,
  chains = 2)

```

Look at the trace and parameter estimates.

```{r}
traceplot(edu_inctxf_fit)
plot(edu_inctxf_fit)
```

### Model 02: ordinal response? ordinal predictor?

### Model 03: survey adjustments?

### Model 04: hierarchical effects?





