---
title: "ESS: Ordinal Regression"
author: "vasa"
date: "7/18/2019"
output: html_document
---

## Setup

```{r setup, cache=TRUE}
library(tidyverse)
library(essurvey)
library(rstan)

source('../scripts/load_survey_data.R')
```

## Simulation

Response: univariate ordinal
Features: univariate continuour with positive effect on ordinal response

We'll simulate a continuous latent variable and then take its quantiles as cut points.

```{r}
simulation_size <- 1000
response_levels <- 5
true_beta <- 2 # mean effect of X on Y


X <- rnorm(simulation_size, 0, 1)
latent_Y <- rnorm(simulation_size, X * true_beta, 1)

# discretize using quantiles of lantent variable
true_cutpoints <- quantile(latent_Y, probs = seq(from = 0, to = 1, length.out = response_levels))

Y <- cut(latent_Y,
         breaks = true_cutpoints,
         ordered_result = TRUE,
         include.lowest = TRUE)

simulation_data <- list('response' = as.integer(Y), 'education' = X)

ggplot(as_tibble(simulation_data), aes(x = X, y = as.integer(Y))) +
  geom_point() +
  geom_abline(slope = true_beta, intercept = 4) + # the middle choice must be X = 0
  theme_bw()

simulation_data[['N']] <- length(simulation_data[['response']])
simulation_data[['K']] <- length(unique(simulation_data[['response']]))

simulation_fit <- stan(file = '../models/ordered_logistic.stan',
  data = simulation_data,
  iter = 2000,
  chains = 1)
```

Take a look if we recover our simulated effect.

```{r}
param_samples <- extract(simulation_fit)

cutpoints_samples <- as_tibble(param_samples[['cutpoints']]) %>%
  rename_all(function(label) str_replace(label, 'V', 'cutpoint_')) %>%
  gather(key = cutpoint, value = sample)

true_cutpoints_data <- as_tibble(true_cutpoints)

ggplot(cutpoints_samples, aes(y = sample, group = cutpoint)) +
  geom_boxplot() + 
  ggtitle('cutpoints') +
  geom_hline(data = true_cutpoints_data, aes(yintercept =  value)) +
  theme_bw()

beta_samples <-  as_tibble(param_samples[['beta']]) %>% rename(beta = value)

true_beta_data <- as_tibble(true_beta)

ggplot(beta_samples, aes(y = beta)) +
  geom_boxplot() + 
  geom_hline(data = true_beta_data, aes(yintercept = value)) +
  ggtitle('beta') +
  theme_bw()
```

## Notes and References

### From the Stan User's Guide

[23.2 Map Function](https://mc-stan.org/docs/2_19/stan-users-guide/map-function.html)

[23.3 Example: Mapping Logistic Regression](https://mc-stan.org/docs/2_19/stan-users-guide/example-mapping-logistic-regression.html)

[23.4 Example: Hierarchical Logistic Regression](https://mc-stan.org/docs/2_19/stan-users-guide/example-hierarchical-logistic-regression.html)

### Forums

["Multilevel Ordered Logistic regression in Stan"](https://stats.stackexchange.com/questions/225963/multilevel-ordered-logistic-regression-in-stan)  

["Stan: Multilevel Ordinal Logistic Regression"](https://groups.google.com/forum/#!category-topic/stan-users/general/sgX2Edo8qiQ)  

["Prediction with ordered logistic regression"](https://discourse.mc-stan.org/t/prediction-with-ordered-logistic-regression/5192)

## Model 01: Basic Ordered Logistic

Start with a simple ordered logistic model.

Taken from Section 1.8 of the 2.18 Stan User's Manual: [Ordered Logistic and Probit Regression](https://mc-stan.org/docs/2_18/stan-users-guide/ordered-logistic-section.html).

```{r}
preprocessed_data <- responses %>%
  select(c(eduyrs, inctxff)) %>%
  drop_na() %>%
  mutate(education = scale(eduyrs) %>% as.vector(), response = as.integer(inctxff) %>% as.vector())

stan_data <- preprocessed_data %>%
  select(c(education, response)) %>%
  as.list()

stan_data[['N']] <- length(stan_data[['response']])
stan_data[['K']] <- length(unique(stan_data[['response']]))

edu_inctxf_fit <-stan(file = '../models/ordered_logistic.stan',
  data = stan_data,
  iter = 2000,
  warmup = 200,
  chains = 2)
```

This takes quite a while to file on my computer.

## Model Checking

We don't even have time for this.

## Model 02: Threaded Ordered Logistic

We can try parallelizing this based on Rickard McElreath's tutorial: [Multithreading and Map-Reduce in Stan 2.18.0: A Minimal Example](https://github.com/rmcelreath/cmdstan_map_rect_tutorial)

The [`map_rect`](https://mc-stan.org/docs/2_19/stan-users-guide/map-function.html) section in the manual also looks good.

```{r}
# Important to run this before compiling the stan model!
# edit_r_makevars()
# CXX14FLAGS = -DSTAN_THREADS -pthread
# CXX14FLAGS += -O3 -march=native -mtune=native
# CXX14FLAGS += -fPIC

Sys.setenv(STAN_NUM_THREADS = 2)

edu_inctxf_fit <-stan(file = '../models/threaded_ordered_logistic.stan',
  data = stan_data,
  iter = 2000,
  warmup = 200,
  chains = 2)
```
