---
title: "ESS: Ordinal Regression"
author: "vasa"
date: "7/18/2019"
output: html_document
---

## Setup

```{r setup, cache=TRUE}
library(tidyverse)
library(essurvey)
library(rstan)

source('../scripts/load_survey_data.R')
```

## Notes and References

### forums

["Multilevel Ordered Logistic regression in Stan"](https://stats.stackexchange.com/questions/225963/multilevel-ordered-logistic-regression-in-stan)  

["Stan: Multilevel Ordinal Logistic Regression"](https://groups.google.com/forum/#!category-topic/stan-users/general/sgX2Edo8qiQ)  

["Prediction with ordered logistic regression"](https://discourse.mc-stan.org/t/prediction-with-ordered-logistic-regression/5192)

## Model 01: Basic Ordered Logistic

Start with a simple ordered logistic model.

Taken from Section 1.8 of the 2.18 Stan User's Manual: [Ordered Logistic and Probit Regression](https://mc-stan.org/docs/2_18/stan-users-guide/ordered-logistic-section.html).

```{r}
preprocessed_data <- responses %>%
  select(c(eduyrs, inctxff)) %>%
  drop_na() %>%
  mutate(education = scale(eduyrs) %>% as.vector(), response = as.integer(inctxff) %>% as.vector())

stan_data <- preprocessed_data %>%
  select(c(education, response)) %>%
  as.list()

stan_data[['N']] <- length(stan_data[['response']])
stan_data[['K']] <- length(unique(stan_data[['response']]))

edu_inctxf_fit <-stan(file = '../models/ordered_logistic.stan',
  data = stan_data,
  iter = 2000,
  warmup = 200,
  chains = 2)
```

This takes quite a while to file on my computer.

## Model Checking

We don't even have time for this.

## Model 02: Threaded Ordered Logistic

We can try parallelizing this based on Rickard McElreath's tutorial: [Multithreading and Map-Reduce in Stan 2.18.0: A Minimal Example](https://github.com/rmcelreath/cmdstan_map_rect_tutorial)

```{r}
# edit_r_makevars()
# CXX14FLAGS = -DSTAN_THREADS -pthread
# CXX14FLAGS += -O3 -march=native -mtune=native
# CXX14FLAGS += -fPIC
Sys.setenv(STAN_NUM_THREADS = 4)

edu_inctxf_fit <-stan(file = '../models/threaded_ordered_logistic.stan',
  data = stan_data,
  iter = 2000,
  warmup = 200,
  chains = 2)
```
